The Dynamics of Normalization Reweighting (target journal: Neuron)

### 2022-06-06 manuscript.Rmd
Update manuscript.Rmd

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','remotes','tictoc','R.matlab','bookdown') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
if (!'osfr' %in% installed.packages()[,1]){remotes::install_github("centerforopenscience/osfr")}
if (!'FourierStats' %in% installed.packages()[,1]){remotes::install_github("bakerdh/FourierStats")}
packagelist <- c(packagelist,'osfr','FourierStats')
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

knitr::opts_chunk$set(echo = TRUE)

@@

targetelectrodes <- c('Oz','POz','O1','O2')
# targetelectrodes <- c('PO8','PO6','PO5','PO7')

@@

@@

      for (tr in 1:counter) {
        cond <- EEGdata$Trigger[triggertimes[tr]] / 10
        condcounter[cond] <- condcounter[cond] + 1
        for (ch in 1:64){
        temp <- as.matrix(EEGdata[(triggertimes[tr]:(triggertimes[tr] + samplerate * duration - 1))-999, ch + 2])
        temp <- temp - mean(temp[1:1000])   # baseline subtraction to remove DC component
        allsamples[condcounter[cond],cond,ch,1:(1000*duration)] <- as.vector(rowMeans(temp))
        }
      }
    }
allsamples[which(is.na(allsamples))] <- 0     # remove any NaN values
meanwaveforms <- array(0,dim=c(14,64,14000))
    for (cond in 1:14){
      for (ch in 1:64){
        temp <- allsamples[1:condcounter[cond],cond,ch,]
        allspec <- matrix(0,nrow=condcounter[cond],ncol=10000)
        for (trial in 1:condcounter[cond]){
          allspec[trial,] <- fft(temp[trial,2001:12000])/10000
        }
        targetresps <- allspec[,tindex]
        maskresps <- allspec[,mindex]
      if (sum(abs(targetresps))>0){  
      tempxy <- data.frame(Re(targetresps),Im(targetresps))
      D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      i1 <- (D<3)
      tempxy <- data.frame(Re(maskresps),Im(maskresps))
      D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      i2 <- (D<3)
      i <- which((i1*i2)>0)
        meanwaveforms[cond,ch,] <- colMeans(temp[i,])
      }
      }
    }

@@

# plot(times,fspec,type='l',col='red')

@@

f1 <- 7
f2 <- 5
    binwidth <- 1000
    tindex <- 1 + f1*(binwidth/1000)
    mindex <- 1 + f2*(binwidth/1000)
condlist <- c(6,8,13)

d <- dir('temp/VilidaiteProcessed/')
all7Hz <- array(0,dim=c(length(d),14,64))
all5Hz <- array(0,dim=c(length(d),14,64))
timecourse7Hz <- array(0,dim=c(length(d),3,14000))
timecourse5Hz <- array(0,dim=c(length(d),3,14000))

  for (s in 1:length(d)){

    load(paste0('temp/VilidaiteProcessed/',d[s]))

    targetchannels <- match(targetelectrodes, electrodes) - 2

    for (cond in 1:14){
      for (ch in 1:64){
        temp <- meanwaveforms[cond,ch,2001:12000]
        spec <- fft(temp)/length(temp)
        all7Hz[s,cond,ch] <- spec[7*10+1]
        all5Hz[s,cond,ch] <- spec[5*10+1]
      }
    }
    for (cond in 1:3){
    temp <- colMeans(meanwaveforms[condlist[cond],targetchannels,])

fspec1 <- temp*0
fspec2 <- temp*0
for (t in 1:(length(temp)-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  fspec1[t] <- spec[tindex]
  fspec2[t] <- spec[mindex]
}
    timecourse7Hz[s,cond,] <- fspec1
    timecourse5Hz[s,cond,] <- fspec2

@@

mean7 <- apply(all7Hz[,,targetchannels],1:2,mean)
mean5 <- apply(all5Hz[,,targetchannels],1:2,mean)
inclusionmatrix7 <- matrix(0,nrow=length(d),ncol=length(condlist))
inclusionmatrix5 <- matrix(0,nrow=length(d),ncol=length(condlist))
for (cond in 1:3){
  temp <- mean7[,condlist[cond]]
        tempxy <- data.frame(Re(temp),Im(temp))
      D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      inclusionmatrix7[,cond] <- (D<3)
   temp <- mean5[,condlist[cond]]
        tempxy <- data.frame(Re(temp),Im(temp))
      D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      inclusionmatrix5[,cond] <- (D<3)
}

meantime5Hz <- matrix(0,nrow=length(condlist),ncol=14000) 
meantime7Hz <- matrix(0,nrow=length(condlist),ncol=14000) 
for (cond in 1:3){
meantime7Hz[cond,] <- abs(apply(abs(timecourse7Hz[which(inclusionmatrix7[,cond]>0),cond,]),2,mean))
meantime5Hz[cond,] <- abs(apply(abs(timecourse5Hz[which(inclusionmatrix5[,cond]>0),cond,]),2,mean))
}

plot(-499:13500,meantime5Hz[2,],type='l',col='red')
lines(-499:13500,meantime5Hz[1,],col='black')
lines(-499:13500,meantime5Hz[3,],col='green')

plot(-499:13500,meantime7Hz[1,],type='l',col='red')
lines(-499:13500,meantime7Hz[2,],col='black')
lines(-499:13500,meantime7Hz[3,],col='green')

@@

```
Our initial reanalysis was promising, however the data were noisy despite the large sample size, because each participant contributed only 8 trials to each condition. We therefore preregistered two new experiments (see http://) to investigate these effects in greater detail. These had a similar overall design to the Vilidaite study, with some small changes intended to optimise the study (see Methods). The key differences were that we used shorter trials (because there were few changes in the latter part of the trials shown in Figure 1), and also focussed all trials into a smaller number of conditions, such that each participant contributed 48 repetitions to each of 4 conditions. The greater data fidelity afforded by this design allowed us to perform a sliding time window analysis of the results.

Figure 2 summarises the results of an EEG experiment testing 100 neurotypical participants. Averaged EEG waveforms showed a strong oscillatory component at each of the two stimulus flicker frequencies (Figure 2a), which slightly lagged the driving signal. Signals were well-isolated in the Fourier domain, and localised to occipital electrodes (Figure 2b). The timecourse at both frequencies showed an initial onset transient, and were then relatively stable for the 6 seconds of stimulus presentation (Figure 2c,d). Responses were weaker in the two masking conditions, and the ratio of target only to target + mask conditions increased over time (Figure 2e,f).

@@

      qdatasubj <- read.csv(paste0('temp/raw/S',subjcounter,'/questionnairedata.csv'))
      save(file=paste0('temp/EEGprocessed/S',subjcounter,'data.RData'),list=c('qdatasubj','condcounter','trialsincluded','subjwaves'))
    }

@@

# d <- dir(path = 'temp/EEGprocessed', pattern = '*.RData', full.names=TRUE)

>>>>

plotsmoothed <- 0 # either plot the raw data (0), or a smoothed spline (1)

# reasonably compact code to check which packages are installed, install the missing ones, and activate all
packagelist <- c('knitr','remotes','tictoc','R.matlab','bookdown','grImport','tiff') # list of CRAN packages
missingpackages <- packagelist[!packagelist %in% installed.packages()[,1]]
if (length(missingpackages)>0){install.packages(missingpackages)}
if (!'osfr' %in% installed.packages()[,1]){remotes::install_github("centerforopenscience/osfr")}
if (!'FourierStats' %in% installed.packages()[,1]){remotes::install_github("bakerdh/FourierStats")}
packagelist <- c(packagelist,'osfr','FourierStats')
toinstall <- packagelist[which(!packagelist %in% (.packages()))]
invisible(lapply(toinstall,library,character.only=TRUE))

@@

v4Interp <- function(df, xo, yo, rmax = .75, gridRes = 67) {
  ## Create a function to perform Matlab's v4 interpolation.
  ## Takes as input a data-frame with columns x, y, and z (x co-ordinates, y co-ordinates, and amplitude)
  ## and variables xo and yo, the co-ordinates which will be use to create a grid for interpolation
  xo <- matrix(rep(xo,length(yo)),nrow = length(xo),ncol = length(yo))
  yo <- t(matrix(rep(yo,length(xo)),nrow = length(yo),ncol = length(xo)))
  xy <- df$x + df$y*sqrt(as.complex(-1))
  d <- matrix(rep(xy,length(xy)),nrow = length(xy), ncol = length(xy))
  d <- abs(d - t(d))
  diag(d) <- 1
  g <- (d^2) * (log(d)-1)   # Green's function.
  diag(g) <- 0
  weights <- qr.solve(g,df$z)
  xy <- t(xy)
  outmat <- matrix(nrow = gridRes,ncol = gridRes)
  for (i in 1:gridRes){
    for (j in 1:gridRes) {
      test4 <- abs((xo[i,j] + sqrt(as.complex(-1))*yo[i,j]) - xy)
      g <- (test4^2) * (log(test4)-1)
      outmat[i,j] <- g %*% weights}}
  outDf <- data.frame(x = xo[,1],outmat)
  names(outDf)[1:length(yo[1,])+1] <- yo[1,]
  return(outDf)}

knitr::opts_chunk$set(echo = TRUE)
colpal <- c('#000000','#8783CF','#FE5000','#004731','#603FB3')

@@

targetelectrodes <- c('Oz','POz','O1','O2')

@@

Our analysis of 220 human M/EEG datasets

@@

      for (tr in 1:counter) {
        cond <- EEGdata$Trigger[triggertimes[tr]] / 10
        condcounter[cond] <- condcounter[cond] + 1
        for (ch in 1:64){
        temp <- as.matrix(EEGdata[(triggertimes[tr]:(triggertimes[tr] + samplerate * duration - 1))-1499, ch + 2])
        temp <- temp - mean(temp[501:1500])   # baseline subtraction to remove DC component
        allsamples[condcounter[cond],cond,ch,1:(1000*duration)] <- as.vector(rowMeans(temp))
        }
      }
    }
allsamples[which(is.na(allsamples))] <- 0     # remove any NaN values
meanwaveforms <- array(0,dim=c(14,64,14000))
    for (cond in 1:14){
      for (ch in 1:64){
        temp <- allsamples[1:condcounter[cond],cond,ch,]
        allspec <- matrix(0,nrow=condcounter[cond],ncol=10000)
        for (trial in 1:condcounter[cond]){
          allspec[trial,] <- fft(temp[trial,2501:12500])/10000
        }
        targetresps <- allspec[,tindex]
        maskresps <- allspec[,mindex]
      if (sum(abs(targetresps))>0){  
      tempxy <- data.frame(Re(targetresps),Im(targetresps))
      D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      i1 <- (D<3)
      tempxy <- data.frame(Re(maskresps),Im(maskresps))
      D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      i2 <- (D<3)
      i <- which((i1*i2)>0)
        meanwaveforms[cond,ch,] <- colMeans(temp[i,])
      }
      }
    }

@@

# plot(times,fspec,type='l',col=colpal[1])

@@

f1 <- 7
f2 <- 5
    binwidth <- 1000
    tindex <- 1 + f1*(binwidth/1000)
    mindex <- 1 + f2*(binwidth/1000)
    tindex2 <- 1 + 2*f1*(binwidth/1000)
    mindex2 <- 1 + 2*f2*(binwidth/1000)

condlist <- c(6,8,13)

d <- dir(path = 'temp/VilidaiteProcessed', pattern = '*.RData', full.names=TRUE)
EEGtimecourse5Hz <- array(0,dim=c(length(d),3,13000))
EEGtimecourse7Hz <- array(0,dim=c(length(d),3,13000))
EEGtimecourse10Hz <- array(0,dim=c(length(d),3,13000))
EEGtimecourse14Hz <- array(0,dim=c(length(d),3,13000))

for (s in 1:length(d)){
  load(d[s])
  targetchannels <- match(targetelectrodes, electrodes) - 2
  for (cond in 1:3){
    temp <- colMeans(meanwaveforms[condlist[cond],targetchannels,],na.rm=TRUE)
    nainds <- which(is.na(temp))   # find occasional NaN values in the waveform
    temp[nainds] <- 0   # replace with zeros

fspec1 <- NULL
fspec2 <- NULL
fspec3 <- NULL
fspec4 <- NULL
for (t in 1:(14000-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  fspec1[t] <- spec[tindex]
  fspec2[t] <- spec[mindex]
  fspec3[t] <- spec[tindex2]
  fspec4[t] <- spec[mindex2]

}
    EEGtimecourse5Hz[s,cond,] <- fspec1
    EEGtimecourse7Hz[s,cond,] <- fspec2
    EEGtimecourse10Hz[s,cond,] <- fspec3
    EEGtimecourse14Hz[s,cond,] <- fspec4

@@

save(file='temp/Pilotsummary.RData',list=c('EEGtimecourse5Hz','EEGtimecourse7Hz','EEGtimecourse10Hz','EEGtimecourse14Hz','electrodes'))

load('temp/Pilotsummary.RData')
EEGmeantime5Hz <- matrix(0,nrow=3,ncol=13000)
EEGmeantime7Hz <- matrix(0,nrow=3,ncol=13000)
EEGmeantime10Hz <- matrix(0,nrow=3,ncol=13000)
EEGmeantime14Hz <- matrix(0,nrow=3,ncol=13000)

for (cond in 1:3){
EEGmeantime5Hz[cond,] <- apply(abs(EEGtimecourse5Hz[,cond,]),2,mean,na.rm=TRUE)
EEGmeantime7Hz[cond,] <- apply(abs(EEGtimecourse7Hz[,cond,]),2,mean,na.rm=TRUE)
EEGmeantime10Hz[cond,] <- apply(abs(EEGtimecourse10Hz[,cond,]),2,mean,na.rm=TRUE)
EEGmeantime14Hz[cond,] <- apply(abs(EEGtimecourse14Hz[,cond,]),2,mean,na.rm=TRUE)
}

EEGmaskindex5 <- colMeans(20*log10(abs(EEGtimecourse5Hz[,1,])/abs(EEGtimecourse5Hz[,3,])))
EEGmaskindex7 <- colMeans(20*log10(abs(EEGtimecourse7Hz[,2,])/abs(EEGtimecourse7Hz[,3,])))
EEGmaskindex10 <- colMeans(20*log10(abs(EEGtimecourse10Hz[,1,])/abs(EEGtimecourse10Hz[,3,])))
EEGmaskindex14 <- colMeans(20*log10(abs(EEGtimecourse14Hz[,2,])/abs(EEGtimecourse14Hz[,3,])))

times <- -499:12500

plot(times,EEGmeantime5Hz[1,],type='l',col=colpal[1],ylim=c(0,1))
lines(times,EEGmeantime5Hz[2,],col='grey')
lines(times,EEGmeantime5Hz[3,],col=colpal[2])
spl5a <- smooth.spline(times,EEGmeantime5Hz[1,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
spl5a <- smooth.spline(times,EEGmeantime5Hz[2,],df=20)
lines(spl5a$x,spl5a$y,col='grey',lwd=2)
spl5a <- smooth.spline(times,EEGmeantime5Hz[3,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[2],lwd=2)

plot(times,EEGmeantime7Hz[2,],type='l',col=colpal[1],ylim=c(0,1))
lines(times,EEGmeantime7Hz[1,],col='grey')
lines(times,EEGmeantime7Hz[3,],col=colpal[2])
spl5a <- smooth.spline(times,EEGmeantime7Hz[2,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
spl5a <- smooth.spline(times,EEGmeantime7Hz[1,],df=20)
lines(spl5a$x,spl5a$y,col='grey',lwd=2)
spl5a <- smooth.spline(times,EEGmeantime7Hz[3,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[2],lwd=2)

plot(times,EEGmaskindex5,type='l',col=colpal[1],ylim=c(0,6))
lines(times,EEGmaskindex7,col=colpal[2])
lines(times,EEGmaskindex10,col=colpal[4])
lines(times,EEGmaskindex14,col=colpal[5])
lines(times,(EEGmaskindex5+EEGmaskindex7+EEGmaskindex10+EEGmaskindex14)/4,col=colpal[3])

spl5a <- smooth.spline(times,EEGmaskindex5,df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
spl5a <- smooth.spline(times,EEGmaskindex7,df=20)
lines(spl5a$x,spl5a$y,col=colpal[2],lwd=2)
spl5a <- smooth.spline(times,(EEGmaskindex5+EEGmaskindex7+EEGmaskindex10+EEGmaskindex14)/4,df=20)
lines(spl5a$x,spl5a$y,col=colpal[3],lwd=4)

# f1 <- 7
# f2 <- 5
#     binwidth <- 1000
#     tindex <- 1 + f1*(binwidth/1000)
#     mindex <- 1 + f2*(binwidth/1000)
# condlist <- c(6,8,13)
# 
# d <- dir('temp/VilidaiteProcessed/')
# all7Hz <- array(0,dim=c(length(d),14,64))
# all5Hz <- array(0,dim=c(length(d),14,64))
# timecourse7Hz <- array(0,dim=c(length(d),3,14000))
# timecourse5Hz <- array(0,dim=c(length(d),3,14000))
# 
#   for (s in 1:length(d)){
# 
#     load(paste0('temp/VilidaiteProcessed/',d[s]))
#     
#     targetchannels <- match(targetelectrodes, electrodes) - 2
# 
#     for (cond in 1:14){
#       for (ch in 1:64){
#         temp <- meanwaveforms[cond,ch,2001:12000]
#         spec <- fft(temp)/length(temp)
#         all7Hz[s,cond,ch] <- spec[7*10+1]
#         all5Hz[s,cond,ch] <- spec[5*10+1]
#       }
#     }
#     for (cond in 1:3){
#     temp <- colMeans(meanwaveforms[condlist[cond],targetchannels,])
# 
# fspec1 <- temp*0
# fspec2 <- temp*0
# for (t in 1:(length(temp)-binwidth)){
#   spec <- fft(temp[t:(t+binwidth-1)])/binwidth
#   fspec1[t] <- spec[tindex]
#   fspec2[t] <- spec[mindex]
# }
#     timecourse7Hz[s,cond,] <- fspec1
#     timecourse5Hz[s,cond,] <- fspec2
#     
#     }
#     
#   }
# 
# mean7 <- apply(all7Hz[,,targetchannels],1:2,mean)
# mean5 <- apply(all5Hz[,,targetchannels],1:2,mean)
# inclusionmatrix7 <- matrix(0,nrow=length(d),ncol=length(condlist))
# inclusionmatrix5 <- matrix(0,nrow=length(d),ncol=length(condlist))
# for (cond in 1:3){
#   temp <- mean7[,condlist[cond]]
#         tempxy <- data.frame(Re(temp),Im(temp))
#       D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
#       inclusionmatrix7[,cond] <- (D<3)
#    temp <- mean5[,condlist[cond]]
#         tempxy <- data.frame(Re(temp),Im(temp))
#       D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
#       inclusionmatrix5[,cond] <- (D<3)
# }
# 
# meantime5Hz <- matrix(0,nrow=length(condlist),ncol=14000) 
# meantime7Hz <- matrix(0,nrow=length(condlist),ncol=14000) 
# for (cond in 1:3){
# meantime7Hz[cond,] <- abs(apply(abs(timecourse7Hz[which(inclusionmatrix7[,cond]>0),cond,]),2,mean))
# meantime5Hz[cond,] <- abs(apply(abs(timecourse5Hz[which(inclusionmatrix5[,cond]>0),cond,]),2,mean))
# }
# 
# plot(-499:13500,meantime5Hz[2,],type='l',col=colpal[1])
# lines(-499:13500,meantime5Hz[1,],col='grey')
# lines(-499:13500,meantime5Hz[3,],col=colpal[2])
# 
# plot(-499:13500,meantime7Hz[1,],type='l',col=colpal[1])
# lines(-499:13500,meantime7Hz[2,],col='grey')
# lines(-499:13500,meantime7Hz[3,],col=colpal[2])
# 
# plot(-499:13500,meantime7Hz[1,]/meantime7Hz[3,],type='l',col=colpal[1])
# lines(-499:13500,meantime5Hz[2,]/meantime5Hz[3,],col=colpal[2])

@@

```
Our initial reanalysis was promising, however the data were noisy despite the large sample size, because each participant contributed only 8 trials to each condition. We therefore preregistered two new experiments (see https://osf.io/4qudc) to investigate these effects in greater detail. These had a similar overall design to the Vilidaite study, with some small changes intended to optimise the study (see Methods). The key differences were that we used shorter trials (because there were few changes in the latter part of the trials shown in Figure 1), and also focussed all trials into a smaller number of conditions, such that each participant contributed 48 repetitions to each of 4 conditions. The greater data fidelity afforded by this design allowed us to perform a sliding time window analysis of the results.

Figure \@ref(fig:EEGdata) summarises the results of an EEG experiment testing 100 neurotypical participants. Averaged EEG waveforms showed a strong oscillatory component at each of the two stimulus flicker frequencies (Figure \@ref(fig:EEGdata)a), which slightly lagged the driving signal. Signals were well-isolated in the Fourier domain, and localised to occipital electrodes (Figure \@ref(fig:EEGdata)b). The timecourse at both frequencies showed an initial onset transient, and was then relatively stable for the 6 seconds of stimulus presentation (Figure \@ref(fig:EEGdata)c,d). Responses were weaker in the two masking conditions, and the ratio of target only to target + mask conditions increased over time (Figure \@ref(fig:EEGdata)e,f) for both mask types. At 5Hz the increase in masking continued over the first 5 seconds of stimulus presentation (Figure \@ref(fig:EEGdata)e), whereas at 7Hz the increase occurred mostly during the first second after onset (Figure \@ref(fig:EEGdata)f).

@@

      qdatasubj <- read.csv(paste0('temp/raw/S',subjcounter,'/questionnairedata.csv'))
      save(file=paste0('temp/EEGprocessed/S',subjcounter,'data.RData'),list=c('qdatasubj','condcounter','trialsincluded','subjwaves','electrodes'))
    }

@@

d <- dir(path = 'temp/EEGprocessed', pattern = '*.RData', full.names=TRUE)
EEGtimecourse5Hz <- array(0,dim=c(length(d),4,8000))
EEGtimecourse7Hz <- array(0,dim=c(length(d),4,8000))
allwaves <- array(0,dim=c(length(d),4,9000))
allspec <- array(0,dim=c(length(d),4,duration*20))
allheads5 <- array(0,dim=c(length(d),4,64))
allheads7 <- array(0,dim=c(length(d),4,64))
for (s in 1:length(d)){
  load(d[s])
  targetchannels <- match(targetelectrodes, electrodes) - 2
  for (cond in 1:4){
    temp <- rowMeans(subjwaves[cond,,targetchannels],na.rm=TRUE)
    nainds <- which(is.na(temp))   # find occasional NaN values in the waveform
    # temp[nainds] <- (temp[nainds-1] + temp[nainds+1])/2   # interpolate adjacent values
    temp[nainds] <- 0   # replace with zeros
    allwaves[s,cond,] <- temp - mean(temp)
    spec <- fft(temp[1501:7500])/6000
    allspec[s,cond,] <- spec[1+(1:(duration*20))]

fspec1 <- NULL
fspec2 <- NULL
for (t in 1:(9000-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  fspec1[t] <- spec[tindex]
  fspec2[t] <- spec[mindex]
}
    EEGtimecourse5Hz[s,cond,] <- fspec1
    EEGtimecourse7Hz[s,cond,] <- fspec2

  for (ch in 1:64){
    temp <- fft(subjwaves[cond,1501:7500,ch])/6000
    allheads5[s,cond,ch] <- temp[1+(duration*targetF)]
    allheads7[s,cond,ch] <- temp[1+(duration*maskF)]
  }
  }

save(file='temp/EEGsummary.RData',list=c('allheads5','allheads7','allspec','allwaves','EEGtimecourse5Hz','EEGtimecourse7Hz','electrodes'))
 }

if (runcode==1){
load('temp/EEGsummary.RData')

isincluded <- matrix(rep(1,4*dim(allspec)[1]),nrow=4,ncol=dim(allspec)[1])
for (cond in 1:4){
  # note that in allspec the DC component has been removed
  # so we don't need to add 1 to the index
      tempxy <- data.frame(Re(allspec[,cond,duration*targetF]),Im(allspec[,cond,duration*targetF]))
      D5 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      tempxy <- data.frame(Re(allspec[,cond,duration*maskF]),Im(allspec[,cond,duration*maskF]))      
      D7 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))      
      isincluded[cond,which(D5>3)] <- 0
      isincluded[cond,which(D7>3)] <- 0
   for (s in 1:100){
     if (max(abs(EEGtimecourse5Hz[s,cond,]))>5){isincluded[cond,s] <- 0}
   }   
}

allwaves[which(allwaves<(-100))] <- 0
allwaves[which(allwaves>100)] <- 0

EEGmeantime5Hz <- matrix(0,nrow=4,ncol=8000)
EEGmeantime7Hz <- matrix(0,nrow=4,ncol=8000)
EEGSEtime5Hz <- matrix(0,nrow=4,ncol=8000)
EEGSEtime7Hz <- matrix(0,nrow=4,ncol=8000)
meanspec <- matrix(0,nrow=4,ncol=120)
meanwave <- matrix(0,nrow=4,ncol=9000)
meanhead5 <- matrix(0,nrow=4,ncol=64)
meanhead7 <- matrix(0,nrow=4,ncol=64)
for (cond in 1:4){
EEGmeantime7Hz[cond,] <- apply(abs(EEGtimecourse7Hz[which(isincluded[cond,]==1),cond,]),2,mean,na.rm=TRUE)
EEGmeantime5Hz[cond,] <- apply(abs(EEGtimecourse5Hz[which(isincluded[cond,]==1),cond,]),2,mean,na.rm=TRUE)

EEGSEtime7Hz[cond,] <- apply(abs(EEGtimecourse7Hz[which(isincluded[cond,]==1),cond,]),2,sd,na.rm=TRUE)/sqrt(sum(isincluded[cond,]))
EEGSEtime5Hz[cond,] <- apply(abs(EEGtimecourse5Hz[which(isincluded[cond,]==1),cond,]),2,sd,na.rm=TRUE)/sqrt(sum(isincluded[cond,]))

meanspec[cond,] <- abs(colMeans(allspec[which(isincluded[cond,]==1),cond,],na.rm=TRUE))
meanwave[cond,] <- colMeans(allwaves[which(isincluded[cond,]==1),cond,],na.rm=TRUE)
meanhead5[cond,] <- abs(colMeans(allheads5[which(isincluded[cond,]==1),cond,],na.rm=TRUE))
meanhead7[cond,] <- abs(colMeans(allheads7[which(isincluded[cond,]==1),cond,],na.rm=TRUE))
}

EEGmaskindex5 <- matrix(0,nrow=2,ncol=8000)
EEGmaskindexSE5 <- matrix(0,nrow=2,ncol=8000)
includedboth <- isincluded[1,]*isincluded[3,]
monratios <- abs(EEGtimecourse5Hz[which(includedboth==1),1,])/abs(EEGtimecourse5Hz[which(includedboth==1),3,])
EEGmaskindex5[1,] <- colMeans(20*log10(monratios))
EEGmaskindexSE5[1,] <- apply(20*log10(monratios),2,sd)/sqrt(sum(includedboth))
includedboth <- isincluded[1,]*isincluded[4,]
dichratios <- abs(EEGtimecourse5Hz[which(includedboth==1),1,])/abs(EEGtimecourse5Hz[which(includedboth==1),4,])
EEGmaskindex5[2,] <- colMeans(20*log10(dichratios))
EEGmaskindexSE5[2,] <- apply(20*log10(dichratios),2,sd)/sqrt(sum(includedboth))

EEGmaskindex7 <- matrix(0,nrow=2,ncol=8000)
EEGmaskindexSE7 <- matrix(0,nrow=2,ncol=8000)
includedboth <- isincluded[2,]*isincluded[3,]
monratios <- abs(EEGtimecourse7Hz[which(includedboth==1),2,])/abs(EEGtimecourse7Hz[which(includedboth==1),3,])
EEGmaskindex7[1,] <- colMeans(20*log10(monratios))
EEGmaskindexSE7[1,] <- apply(20*log10(monratios),2,sd)/sqrt(sum(includedboth))
includedboth <- isincluded[2,]*isincluded[4,]
dichratios <- abs(EEGtimecourse7Hz[which(includedboth==1),2,])/abs(EEGtimecourse7Hz[which(includedboth==1),4,])
EEGmaskindex7[2,] <- colMeans(20*log10(dichratios))
EEGmaskindexSE7[2,] <- apply(20*log10(dichratios),2,sd)/sqrt(sum(includedboth))

# pdf("temp/figures/Fig3.pdf", bg="transparent", height = 10, width = 12)
# 
# par(mfrow=c(3,4))

times <- ((1:8000)-1000)/1000

postscript("timecourse1.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5, width = 5)

plotlims <- c(-1,7,0,0.6)
ticklocsx <- -1:7    # locations of tick marks on x axis
ticklocsy <- seq(0,0.6,0.1)    # locations of tick marks on y axis
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocsx, side = 1, at=ticklocsx)
mtext(text = ticklocsy, side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Amplitude (µV)", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

polygon(c(0,6,6,0),c(-0.022,-0.022,0.02,0.02),border=NA,col=rgb(0.5,0.5,0.5))

if (plotsmoothed==0){
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[1,]+EEGSEtime5Hz[1,],EEGmeantime5Hz[1,8000:1]-EEGSEtime5Hz[1,8000:1]),col=colpal[1],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[3,]+EEGSEtime5Hz[3,],EEGmeantime5Hz[3,8000:1]-EEGSEtime5Hz[3,8000:1]),col=colpal[2],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[4,]+EEGSEtime5Hz[4,],EEGmeantime5Hz[4,8000:1]-EEGSEtime5Hz[4,8000:1]),col=colpal[3],border=NA)

lines(times,EEGmeantime5Hz[1,],lwd=2,col=colpal[1])
# lines(times,EEGmeantime5Hz[2,],lwd=2,col='green')
lines(times,EEGmeantime5Hz[3,],lwd=2,col=colpal[2])
lines(times,EEGmeantime5Hz[4,],lwd=2,col=colpal[3])
}

if (plotsmoothed==1){
spl5aU <- smooth.spline(times,EEGmeantime5Hz[1,]+EEGSEtime5Hz[1,],df=20)
spl5aL <- smooth.spline(times,EEGmeantime5Hz[1,]-EEGSEtime5Hz[1,],df=20)
polygon(times[c(1:8000,8000:1)],c(spl5aU$y,spl5aL$y[8000:1]),col=colpal[1],border=NA)

spl5cU <- smooth.spline(times,EEGmeantime5Hz[3,]+EEGSEtime5Hz[3,],df=20)
spl5cL <- smooth.spline(times,EEGmeantime5Hz[3,]-EEGSEtime5Hz[3,],df=20)
polygon(times[c(1:8000,8000:1)],c(spl5cU$y,spl5cL$y[8000:1]),col=colpal[2],border=NA)

spl5dU <- smooth.spline(times,EEGmeantime5Hz[4,]+EEGSEtime5Hz[4,],df=20)
spl5dL <- smooth.spline(times,EEGmeantime5Hz[4,]-EEGSEtime5Hz[4,],df=20)
polygon(times[c(1:8000,8000:1)],c(spl5dU$y,spl5dL$y[8000:1]),col=colpal[3],border=NA)

spl5a <- smooth.spline(times,EEGmeantime5Hz[1,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
# spl5b <- smooth.spline(times,EEGmeantime5Hz[2,],df=20)
# lines(spl5b$x,spl5b$y,col='green',lwd=2)
spl5c <- smooth.spline(times,EEGmeantime5Hz[3,],df=20)
lines(spl5c$x,spl5c$y,col=colpal[2],lwd=2)
spl5d <- smooth.spline(times,EEGmeantime5Hz[4,],df=20)
lines(spl5d$x,spl5d$y,col=colpal[3],lwd=2)

dev.off()

postscript("timecourse2.ps", horizontal = FALSE, onefile = FALSE, paper = "special", height = 5, width = 5)

plotlims <- c(-1,7,-3,7)
ticklocsy <- seq(-3,6,3)    # locations of tick marks on y axis
plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=plotlims[1:2], ylim=plotlims[3:4])
axis(1, at=ticklocsx, tck=0.01, lab=F, lwd=2)
axis(2, at=ticklocsy, tck=0.01, lab=F, lwd=2)
mtext(text = ticklocsx, side = 1, at=ticklocsx)
mtext(text = c(expression(sqrt(2)/2),'1',expression(sqrt(2)),'2'), side = 2, at=ticklocsy, line=0.2, las=1)
title(xlab="Time (s)", col.lab=rgb(0,0,0), line=1.2, cex.lab=1.5)    
title(ylab="Suppression ratio", col.lab=rgb(0,0,0), line=1.5, cex.lab=1.5)

polygon(c(0,6,6,0),c(-3.4,-3.4,-2.6,-2.6),border=NA,col=rgb(0.5,0.5,0.5))

lines(c(-1,7),c(0,0),lty=2)

if (plotsmoothed==0){
polygon(times[c(1:8000,8000:1)],c(EEGmaskindex5[1,]+EEGmaskindexSE5[1,],EEGmaskindex5[1,8000:1]-EEGmaskindexSE5[1,8000:1]),col=colpal[2],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmaskindex5[2,]+EEGmaskindexSE5[2,],EEGmaskindex5[2,8000:1]-EEGmaskindexSE5[2,8000:1]),col=colpal[3],border=NA)
